---
title: "MECS 5.4 Clean - Disaggregated"
author: "Jaxon Stuhr"
date: "1/27/2022"
output: html_document
---

This script cleans the MECS 5.4 table and disaggregates all sectos. This table provides fuel consumption by end use and fuel type for all NAICS manufacturing sectors (all units = TBtu). NAICS categories are "disaggregated", meaning 3-digit and 4-digit NAICS codes DO NOT include energy consumption of their subsectors. These should be thought of as "3xx (other)" This dataset can be summed over sectors (NAICS codes) to get total fuel consumption, as there will be no double counting of subsector consumption. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(writexl)
```

```{r}
# read in raw mecs tables
raw5_3 = read_excel(here("data", "Table5_3.xlsx")) # UNITS = varied, million kWh or million bbl or billion cu ft or million short tons
raw5_4 = read_excel(here("data", "Table5_4.xlsx")) # UNITS = Tbtu
```

```{r}
# clean data 
mecs5_4 = raw5_4 %>% 
  # delete top rows
  tail(-14) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "end_use" = "...2",
         "net_electricity_demand" = "...3",
         "residual_fuel_oil" = "...4",
         "distillate_fuel_oil_and_diesel_fuel" = "...5",
         "nat_gas" = "...6",
         "hgl_excluding_ng" = "...7",
         "coal_excluding_coke_and_breeze" = "...8") %>%
    # remove cols after end
  select(naics_code : coal_excluding_coke_and_breeze) %>% 
  # remove blank rows
  remove_empty("rows") 
```

```{r}
# set NAICS code vals for all entries with blank codes
counter = 1
for (entry in mecs5_4$naics_code) {
  if (is.na(entry)) {
    mecs5_4$naics_code[counter] = mecs5_4$naics_code[counter - 1]
  }
  counter = counter + 1
}  
```

```{r}
# convert data type to numeric for columns 3 - 10
cols_to_numeric = c(3:8) 
mecs5_4[ , cols_to_numeric] = apply(mecs5_4[ , cols_to_numeric], 2,  # Specify own function within apply
                    function(x) as.numeric(as.character(x)
                                           )
                    )
# convert NAs to 0
mecs5_4[is.na(mecs5_4)] = 0
```

```{#r}
# update all non-numeral values in mecs table
all_end_uses = unique(mecs5_2$end_use[2:19])
indirect_uses = all_end_uses[3:4]
direct_uses_process = all_end_uses[6:10]
direct_uses_non_process = all_end_uses[12:17]
aggregate_end_uses = c(all_end_uses[2],all_end_uses[5],all_end_uses[11],all_end_uses[18])

for (row in 1:nrow(mecs5_2) ) {
  code = naics_code[row]
  end_use = 
  for (col in colnames(mecs5_2)) {
    if ()
  }
  
}
# mecs5_2 = mecs5_2 %>% 
#   mutate(net_electricity = case_when(
#     net_electricity == NA ~ case_when(
#       net_electricity %in% indirect_uses ~ net_electricity
#     )
#   ))
```


```{r}
### Remove all unecessary rows

# remove subtotal and total rows 
mecs5_4 = mecs5_4 %>% 
  filter(!(end_use %in% c("Indirect Uses-Boiler Fuel", 
                          "Direct Uses-Total Process", 
                          "Direct Uses-Total Nonprocess")))
# remove total rows in beginning (naics_code = 311-339)
mecs5_4 = mecs5_4 %>% 
  filter(naics_code != "311 - 339")
# remove scetor label rows
mecs5_4 = mecs5_4 %>% 
  filter(end_use %in% mecs5_4$end_use[2:16]) 
```


```{r}
# function to update aggregated values in MECS table
update_naics_vals = function(new_vals, mecs_table, current_code, next_code, next_row, end_use_count, char_count) {
    # update values of usage columns when match is found (indices are for values that need updating)
    new_vals[3:8] = new_vals[3:8] - mecs_table[next_row, 3:8]
    # update next row (w same end use) and code
    next_row = next_row + end_use_count
    next_code = mecs_table$naics_code[next_row]
    # check if next naics code is also sub-code of current code
    if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
      # if yes, recall update naics function
        return(update_naics_vals(new_vals, mecs_table, current_code, next_code, next_row, end_use_count, char_count))
    } else {
      # if no, return new vals
        return(new_vals)
      }
}
```

## Update 4-digit NAICS Codes

```{r}
# initialize new table
new_mecs_table = mecs5_4[0, ]
# count number of end uses /// number each time each naics code appears
end_use_count = sum(mecs5_4$naics_code == mecs5_4$naics_code[1])
char_count = 4 # updating 4 digit naics codes
# iterate through rows in mecs table
for (row in 1: (nrow(mecs5_4) - end_use_count) ) {
  # define current code, next code, and next row index with same end use as current
  current_code = mecs5_4$naics_code[row]
  next_row = row + end_use_count
  next_code = mecs5_4$naics_code[next_row]
  # build array of vals for row to be updated
  new_vals = mecs5_4[row, ]
  # check next row is subcode of current code
  if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
    # if yes, update vals
          new_mecs_table[nrow(new_mecs_table) + 1,] = 
    update_naics_vals(new_vals, mecs5_4, current_code, next_code, next_row, end_use_count, char_count)
  } else {
    # if not, append row to data
    new_mecs_table[nrow(new_mecs_table) + 1,] = new_vals
  }
}

# add last naics code values on to new data frame
end_of_old_table = mecs5_4[as.numeric(count(mecs5_4) - end_use_count + 1):as.numeric(count(mecs5_4)), ]
new_mecs_table = rbind(new_mecs_table, end_of_old_table)
```

## Update 3-digit NAICS Codes

```{r}
# initialize new table
final_mecs_table = new_mecs_table[0, ]
# count number of end uses /// number each time each naics code appears
end_use_count = sum(new_mecs_table$naics_code == new_mecs_table$naics_code[1])
char_count = 3 # updating 3 digit naics codes
# iterate through rows in mecs table
for (row in 1:(nrow(new_mecs_table) - end_use_count) ) {
  # define current code, next code, and next row index with same end use as current
  current_code = new_mecs_table$naics_code[row]
  next_row = row + end_use_count
  next_code = new_mecs_table$naics_code[next_row]
  # build array of vals for row to be updated
  new_vals = new_mecs_table[row, ]
  # check next row is subcode of current code
  if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
    # if yes, update vals
          final_mecs_table[nrow(final_mecs_table) + 1,] = 
    update_naics_vals(new_vals, new_mecs_table, current_code, next_code, next_row, end_use_count, char_count)
  } else {
    # if not, append row to data
    final_mecs_table[nrow(final_mecs_table) + 1,] = new_vals
  }
}

# add last naics code values on to new data frame
end_of_old_table = new_mecs_table[as.numeric(count(new_mecs_table) - end_use_count + 1) :
                                    as.numeric(count(new_mecs_table)), ]
final_mecs_table = rbind(final_mecs_table, end_of_old_table)
```

```{r}
# write final dataframe to an excel file
mecs5_4_wrangled = final_mecs_table
write_xlsx(mecs5_4_wrangled, here("outputs", "mecs5_4_wrangled.xlsx"))
```

```{#r}
# write second MECS file for steam analysis, only boilers and summing
steam_mecs_5_4 = final_mecs_table %>% 
  filter(end_use == "Conventional Boiler Use") %>% 
  pivot_longer(3:8, names_to = "fuel_type", values_to = "fuel_consumption_tbtu")
```



