---
title: "MECS Byproducts and Wood Fuels Clean - Aggregated"
author: "Jaxon Stuhr"
date: "5/21/2022"
output: html_document
---

This script cleans MECS tables 3.5 (byproduct fuels) and 3.6 (wood fuels) which provide fuel consumption by fuel type for all NAICS manufacturing sectors (all units = TBtu). NAICS categories are "aggregated", meaning 3-digit NAICS codes include all fuel consumption of 4, 5, and 6-digit subsector NAICS codes. This means that the dataset cannot be summed over sectors (NAICS codes) to get total fuel consumption, as there would be double counting of subsector consumption.

This code requires "mecs_5_4_clean_aggregated.Rmd" to be run first, in this environment

Load in libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(writexl)
```

Read in Mecs 3.2, 3.5, 3.6
```{r}
# read in raw mecs tables
raw3_2 = read_excel(here("data", "Table3_2.xlsx")) # UNITS = Tbtu
raw3_5 = read_excel(here("data", "Table3_5.xlsx")) # UNITS = Tbtu
raw3_6 = read_excel(here("data", "Table3_6.xlsx")) # UNITS = Tbtu
```

Rename columns as fuel types and remove filler rows (3.2)
```{r}
# clean table 3.5
# Total united states only
mecs3_2 = raw3_2 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "other" = "...11") %>%
  # remove subsector description column
  select(naics_code, other) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
```

Rename columns as fuel types and remove filler rows (3.5)
```{r}
# clean table 3.5
# Total united states only
mecs3_5 = raw3_5 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "blast_furnace_coke_oven_gases" = "...4",
         "waste_gas" = "...5",
         "petroleum_coke" = "...6",
         "pulping_liquor_or_black_liquor" = "...7",
         "wood_chips_bark" = "...8",
         "waste_oils_tars_waste_minerals" = "...9") %>%
  # remove subsector description column
  select(naics_code, blast_furnace_coke_oven_gases : waste_oils_tars_waste_minerals) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
```

Rename columns as fuel types and remove filler rows (3.6). From 3.6, only ag waste is needed as wood chips from 3.5 covers the other woody biomass included in 3.6.
```{r}
# clean table 3.6
# Total united states only
mecs3_6 = raw3_6 %>% 
  # delete top rows
  tail(-17) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "agricultural_waste" = "...5") %>% 
  # remove subsector description column
  select(naics_code, agricultural_waste) %>%
  # remove blank rows
  remove_empty("rows") %>%
  # total US only
  head(14) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
  # add missing rows with 0s
 missing_codes = mecs3_5$naics_code[!(mecs3_5$naics_code %in% mecs3_6$naics_code)]
  for (code in missing_codes) {
    mecs3_6[nrow(mecs3_6) + 1,] = as.list(c(code,0))
  }
  ### 311221 added but 3112 set to zero, MANUALLY UPDATE 3112!!!
  mecs3_6[15,2] = mecs3_6[2,2]
```

Merge Tables 3.2, 3.5, 3.6 together into "Others". This includes only some of the "Other" fuel types denoted in MECS 5.2
```{r}
# merge all tables by naics code
mecs_others = mecs3_5 %>% 
  merge(mecs3_6, by = "naics_code", all = TRUE, sort = FALSE) %>% 
  merge(mecs3_2, by = "naics_code", all = TRUE, sort = FALSE) %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) # set all NAs to zero
```

Add in rows for end-uses of MECS 5.4, as 3.5 and 3.6 only have sector-level data. This code loops through each NAICS code, adding 15 new rows after each "Total" row for each end use. These values are left blank for now and are filled-in in Excel. 
```{r}
mecs_others = mecs_others %>% 
  mutate(end_use = "TOTAL FUEL CONSUMPTION", .after = naics_code)
end_uses = unique(mecs5_4$end_use)
naics_codes = unique(mecs_others$naics_code)
row = 1
for (code in naics_codes) {
  mecs_others = mecs_others %>% 
    add_row(naics_code = code, end_use = "Conventional Boiler Use", .after = row) %>% 
    add_row(naics_code = code, end_use = "CHP and/or Cogeneration Process", .after = row+1) %>% 
    add_row(naics_code = code, end_use = "Process Heating", .after = row+2) %>% 
    add_row(naics_code = code, end_use = "Process Cooling and Refrigeration", .after = row+3) %>% 
    add_row(naics_code = code, end_use = "Machine Drive", .after = row+4) %>% 
    add_row(naics_code = code, end_use = "Electro-Chemical Processes", .after = row+5) %>% 
    add_row(naics_code = code, end_use = "Other Process Use", .after = row+6) %>% 
    add_row(naics_code = code, end_use = "Facility HVAC (f)", .after = row+7) %>% 
    add_row(naics_code = code, end_use = "Facility Lighting", .after = row+8) %>% 
    add_row(naics_code = code, end_use = "Other Facility Support", .after = row+9) %>% 
    add_row(naics_code = code, end_use = "Onsite Transportation", .after = row+10) %>% 
    add_row(naics_code = code, end_use = "Conventional Electricity Generation", .after = row+11) %>% 
    add_row(naics_code = code, end_use = "Other Nonprocess Use", .after = row+12) %>% 
    add_row(naics_code = code, end_use = "End Use Not Reported", .after = row+13) 
  row = row + 15 #length(end_uses)
}
```

Update "Other" Column to be leftovers after byproducts and biomass
```{r}
mecs_others = mecs_others %>% 
  mutate(other = other - blast_furnace_coke_oven_gases - waste_gas - petroleum_coke - pulping_liquor_or_black_liquor - wood_chips_bark - waste_oils_tars_waste_minerals - agricultural_waste  )
```


Write to excel file
```{r}
write_xlsx(mecs_others, here("outputs", "MECS_Others_Clean.xlsx"))
```

