---
title: "MECS 5.4 Clean - Aggregated"
author: "Jaxon Stuhr"
date: "5/21/2022"
output: html_document
---

This script cleans the MECS 5.4 table. This table provides fuel consumption by end use and fuel type for all NAICS manufacturing sectors (all units = TBtu). NAICS categories are "aggregated", meaning 3-digit NAICS codes include all fuel consumption of 4, 5, and 6-digit subsector NAICS codes. This means that the dataset cannot be summed over sectors (NAICS codes) to get total fuel consumption, as there would be double counting of subsector consumption.

Load in libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(writexl)
```

Read in MECS 5.4, all units are TBtu
```{r}
# read in raw mecs tables
raw5_4 = read_excel(here("data", "Table5_4.xlsx")) # UNITS = Tbtu
```

Rename columns as fuel types and remove filler rows
```{r}
# clean data 
mecs5_4 = raw5_4 %>% 
  # delete top rows
  tail(-14) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "end_use" = "...2",
         "net_electricity_demand" = "...3",
         "residual_fuel_oil" = "...4",
         "distillate_fuel_oil_and_diesel_fuel" = "...5",
         "nat_gas" = "...6",
         "hgl_excluding_ng" = "...7",
         "coal_excluding_coke_and_breeze" = "...8") %>%
    # remove cols after end
  select(naics_code : coal_excluding_coke_and_breeze) %>% 
  # remove blank rows
  remove_empty("rows") 
```

Set correct NAICS code for all rows
```{r}
counter = 1
for (entry in mecs5_4$naics_code) {
  if (is.na(entry)) {
    mecs5_4$naics_code[counter] = mecs5_4$naics_code[counter - 1]
  }
  counter = counter + 1
}  
```

Convert type to numeric for "value" rows
```{r}
cols_to_numeric = c(3:8) 
mecs5_4[ , cols_to_numeric] = apply(mecs5_4[ , cols_to_numeric], 2,  # Specify own function within apply
                    function(x) as.numeric(as.character(x)
                                           )
                    )
# convert NAs to 0
mecs5_4[is.na(mecs5_4)] = 0
```

Remove other unnecessary rows
```{r}
mecs5_4 = mecs5_4 %>% 
  filter(!(end_use %in% c("Indirect Uses-Boiler Fuel", # "subtotal" rows
                          "Direct Uses-Total Process", 
                          "Direct Uses-Total Nonprocess")),
         naics_code != "311 - 339") 
mecs5_4 = mecs5_4 %>% filter(end_use %in% mecs5_4$end_use[2:16]) # other rows
```

Write dataset to excel file
```{r}
# write final dataframe to an excel file
write_xlsx(mecs5_4, here("outputs", "MECS_5_4_Clean.xlsx"))
```




