---
title: "MECS Others Clean - Disaggregated"
author: "Jaxon Stuhr"
date: "2/9/2022"
output: html_document
---

This script cleans MECS tables 3.2, 3.5 (byproduct fuels) and 3.6 (wood fuels), 7.7, 11.1, 11.3, 11.5, which provide fuel consumption by fuel type for all NAICS manufacturing sectors (all units = TBtu). NAICS categories are "disaggregated", meaning 3-digit and 4-digit NAICS codes DO NOT include energy consumption of their subsectors. These should be thought of as "3xx (other)" This dataset can be summed over sectors (NAICS codes) to get total fuel consumption, as there will be no double counting of subsector consumption. Additionally, blank rows are added so that the consumption of "Other" fuels can be disaggregated be end-use (this is performed seperateley in excel).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(writexl)
```

```{r}
# unit conversions
kwh_to_btu = 3412.1416
mil_kwh_to_tbtu = kwh_to_btu * (10^6) / (10^12)
ng_bil_cu_ft_to_tbtu = 1.0388 # calculate from mecs tables 5.3, 5.4
bil_btu_to_tbtu = 0.001
```

```{r}
# read in raw mecs tables
raw3_2 = read_excel(here("data", "Table3_2.xlsx"))
raw3_5 = read_excel(here("data", "Table3_5.xlsx")) # UNITS = Tbtu
raw3_6 = read_excel(here("data", "Table3_6.xlsx")) # UNITS = Tbtu
raw7_7 = read_excel(here("data", "Table7_7.xlsx")) # UNITS = varied, million kWh or billion cu ft or billion btu
raw11_1 = read_excel(here("data", "Table11_1.xlsx")) # UNITS = million kWh
raw11_3 = read_excel(here("data", "Table11_3.xlsx")) # UNITS = million kWH (?)
raw11_5 = read_excel(here("data", "Table11_5.xlsx")) # UNITS = million kWh
```

```{r}
# clean table 3.5
# Total united states only
mecs3_2 = raw3_2 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "coke_and_breeze" = "...10",
         "other" = "...11") %>%
  # remove subsector description column
  select(naics_code, coke_and_breeze, other) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
```

```{r}
# clean table 3.5
# Total united states only
mecs3_5 = raw3_5 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "subsector_and_industry" = "...2",
         "selected_byproducts_in_fuel_consumption_total" = "...3",
         "blast_furnace_coke_oven_gases" = "...4",
         "waste_gas" = "...5",
         "petroleum_coke" = "...6",
         "pulping_liquor_or_black_liquor" = "...7",
         "wood_chips_bark" = "...8",
         "waste_oils_tars_waste_minerals" = "...9") %>%
  # remove subsector description column
  select(naics_code, selected_byproducts_in_fuel_consumption_total : waste_oils_tars_waste_minerals) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
```

```{r}
# clean table 3.6
# Total united states only
mecs3_6 = raw3_6 %>% 
  # delete top rows
  tail(-17) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "pulping_liquor_or_black_liquor" = "...3",
         "biomass_total" = "...4",
         "agricultural_waste" = "...5",
         "wood_harvested_directly_from_trees" = "...6",
         "wood_residues_byproducts_from_mill_processing" = "...7",
         "wood_related_paper_related_refuse" = "...8") %>%
  # remove subsector description column
  select(naics_code, pulping_liquor_or_black_liquor : wood_related_paper_related_refuse) %>%
  # remove blank rows
  remove_empty("rows") %>%
  # total US only
  head(14) %>% 
  # convert values to numeric
  mutate_all(as.numeric)
  # add missing rows with 0s
  missing_codes = mecs3_5$naics_code[!(mecs3_5$naics_code %in% mecs3_6$naics_code)]
  for (code in missing_codes) {
    mecs3_6[nrow(mecs3_6) + 1,] = as.list(c(code,0,0,0,0,0,0))
  }
  ### 311221 added but 3112 set to zero, MANUALLY UPDATE 3112!!!
  mecs3_6[15,2:6] = mecs3_6[2,2:6]
```

```{r}
# clean table 7.7
# Total united states only
mecs7_7 = raw7_7 %>% 
  # delete top rows
  tail(-16) %>% 
  # rename cols
  rename("naics_code" = "Released: September 2021",
         "elec_tot_mil_kwh" = "...3",
         "elec_loc_utility_mil_kwh" = "...4",
         "elec_other_sources_mil_kwh" = "...5",
         "ng_tot_bil_cu_ft" = "...6",
         "ng_loc_utility_bil_cu_ft" = "...7",
         "ng_other_sources_bil_cu_ft" = "...8",
         "steam_tot_bil_btu" = "...9",
         "steam_loc_utility_bil_btu" = "...10",
         "steam_other_sources_bil_btu" = "...11") %>%
  # remove subsector description column
  select(naics_code, elec_tot_mil_kwh : steam_other_sources_bil_btu) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric) %>% 
  # update electricity units to tbtu
  mutate_at(vars(elec_tot_mil_kwh:elec_other_sources_mil_kwh),
            .funs = funs(. * mil_kwh_to_tbtu)) %>%
  # update ng units to tbtu
  mutate_at(vars(ng_tot_bil_cu_ft:ng_other_sources_bil_cu_ft),
            .funs = funs(. * ng_bil_cu_ft_to_tbtu)) %>%
  # update steam units to tbtu
  mutate_at(vars(steam_tot_bil_btu:steam_other_sources_bil_btu),
            .funs = funs(. * bil_btu_to_tbtu)) %>% 
  mutate_all(round)
```

```{r}
# clean table 11.1
# Total united states only
mecs11_1 = raw11_1 %>% 
  # delete top rows
  tail(-11) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "elec_purchases" = "...3",
         "elec_transfers_in" = "...4",
         "elec_total_onsite_generation" = "...5",
         "elec_sales_and_transfers_offsite" = "...6",
         "elec_net_demand_for_electricity" = "...7") %>%
  # remove subsector description column
  select(naics_code, elec_purchases : elec_net_demand_for_electricity) %>%
  # remove blank rows
  remove_empty("rows") %>%
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric) %>% 
  # convert all values from million kWh to TBtu
  mutate_at(vars(elec_purchases:elec_net_demand_for_electricity),
            .funs = funs(. * mil_kwh_to_tbtu)) %>% 
  mutate_all(round)
```

```{r}
# clean table 11.3
# Total united states only
mecs11_3 = raw11_3 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols, drop onsite generation bc included in 11.1
  rename("naics_code" = "Released: February 2021",
         "elec_cogeneration" = "...4",
         "elec_renewable_energy_excluding_wood_and_biomass" = "...5",
         "elec_other" = "...6") %>%
  # remove subsector description column
  select(naics_code, elec_cogeneration : elec_other) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82) %>% 
  # convert values to numeric
  mutate_all(as.numeric) %>% 
  # convert all values from million kWh to TBtu
  mutate_at(vars(elec_cogeneration : elec_other),
            .funs = funs(. * mil_kwh_to_tbtu)) %>% 
  mutate_all(round)
```

```{r}
# clean table 11.3
# Total united states only
mecs11_5 = raw11_5 %>% 
  # delete top rows
  tail(-12) %>% 
  # rename cols
  rename("naics_code" = "Released: February 2021",
         "elec_totals_sales_transfers_offsite" = "...3",
         "elec_utility_purchaser" = "...4",
         "elec_nonutility_purchaser" = "...5") %>%
  # remove subsector description column
  select(naics_code, elec_totals_sales_transfers_offsite : elec_nonutility_purchaser) %>%
  # remove blank rows
  remove_empty("rows") %>% 
  # total US only
  head(82)  %>% 
  # convert values to numeric
  mutate_all(as.numeric) %>% 
  # convert all values from million kWh to TBtu
  mutate_at(vars(elec_totals_sales_transfers_offsite:elec_nonutility_purchaser),
            .funs = funs(. * mil_kwh_to_tbtu)) %>% 
  mutate_all(round)
```

```{r}
# merge all tables by naics code
all_coeffs = mecs3_5 %>% 
  merge(mecs3_6, by = "naics_code", all = TRUE, sort = FALSE) %>% 
  merge(mecs7_7, by = "naics_code", sort = FALSE) %>% 
  merge(mecs11_1, by = "naics_code", sort = FALSE) %>% 
  merge(mecs11_3, by = "naics_code", sort = FALSE) %>% 
  merge(mecs11_5, by = "naics_code", sort = FALSE) %>% 
  merge(mecs3_2, by = "naics_code", sort = FALSE) %>% 
  # set all NAs to zero
  mutate_all(funs(ifelse(is.na(.), 0, .)))
```

```{r}
# function to update aggregated values in MECS table
update_coeffs = function(new_vals, mecs_table, current_code, next_code, next_row, char_count) {
    # update values of usage columns when match is found (indices are for values that need updating)
    new_vals[2:36] = new_vals[2:36] - mecs_table[next_row, 2:36]
    # update next row (w same end use) and code
    next_row = next_row + 1
    next_code = mecs_table$naics_code[next_row]
    # check if next naics code is also sub-code of current code
    if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
      # if yes, recall update naics function
        return(update_coeffs(new_vals, mecs_table, current_code, next_code, next_row, char_count))
    } else {
      # if no, return new vals
        return(new_vals)
      }
}
```

## Update 4-digit NAICS Codes

```{r}
# initialize new table
new_coeffs_4_digit_updated = all_coeffs[0, ]
# updating 4 digit naics codes
char_count = 4 
# iterate through rows in mecs table
for (row in 1: (nrow(all_coeffs) - 1) ) {
  # define current code, next code, and next row index with same end use as current
  current_code = all_coeffs$naics_code[row]
  next_row = row + 1
  next_code = all_coeffs$naics_code[next_row]
  # build array of vals for row to be updated
  new_vals = all_coeffs[row, ]
  # check next row is subcode of current code
  if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
    # if yes, update vals
          new_coeffs_4_digit_updated[nrow(new_coeffs_4_digit_updated) + 1,] = 
            update_coeffs(new_vals, all_coeffs, current_code, next_code, next_row, char_count)
  } else {
    # if not, append row to data
          new_coeffs_4_digit_updated[nrow(new_coeffs_4_digit_updated) + 1,] = new_vals
  }
}
# append last row
new_coeffs_4_digit_updated = rbind(new_coeffs_4_digit_updated, tail(all_coeffs, 1))
```

## Update 3-digit NAICS Codes

```{r}
# initialize new table
final_coeffs_3_digit_updated = all_coeffs[0, ]
# updating 4 digit naics codes
char_count = 3 
# iterate through rows in mecs table
for (row in 1: (nrow(new_coeffs_4_digit_updated) - 1) ) {
  # define current code, next code, and next row index with same end use as current
  current_code = new_coeffs_4_digit_updated$naics_code[row]
  next_row = row + 1
  next_code = new_coeffs_4_digit_updated$naics_code[next_row]
  # build array of vals for row to be updated
  new_vals = new_coeffs_4_digit_updated[row, ]
  # check next row is subcode of current code
  if (
      nchar(current_code) == char_count & 
      nchar(next_code) > char_count & 
      substr(next_code, 1, char_count) == current_code
      ) {
    # if yes, update vals
          final_coeffs_3_digit_updated[nrow(final_coeffs_3_digit_updated) + 1,] = 
            update_coeffs(new_vals, new_coeffs_4_digit_updated, current_code, next_code, next_row, char_count)
  } else {
    # if not, append row to data
          final_coeffs_3_digit_updated[nrow(final_coeffs_3_digit_updated) + 1,] = new_vals
  }
}
# append last row
final_coeffs_3_digit_updated = rbind(final_coeffs_3_digit_updated, tail(new_coeffs_4_digit_updated, 1))
```

```{r}
full_coeffs = final_coeffs_3_digit_updated %>% 
  mutate(end_use = "TOTAL FUEL CONSUMPTION", .after = naics_code)
end_uses = unique(mecs5_4_wrangled$end_use)
naics_codes = unique(final_coeffs_3_digit_updated$naics_code)
row = 1
for (code in naics_codes) {
  full_coeffs = full_coeffs %>% 
    add_row(naics_code = code, end_use = "Conventional Boiler Use", .after = row) %>% 
    add_row(naics_code = code, end_use = "CHP and/or Cogeneration Process", .after = row+1) %>% 
    add_row(naics_code = code, end_use = "Process Heating", .after = row+2) %>% 
    add_row(naics_code = code, end_use = "Process Cooling and Refrigeration", .after = row+3) %>% 
    add_row(naics_code = code, end_use = "Machine Drive", .after = row+4) %>% 
    add_row(naics_code = code, end_use = "Electro-Chemical Processes", .after = row+5) %>% 
    add_row(naics_code = code, end_use = "Other Process Use", .after = row+6) %>% 
    add_row(naics_code = code, end_use = "Facility HVAC (f)", .after = row+7) %>% 
    add_row(naics_code = code, end_use = "Facility Lighting", .after = row+8) %>% 
    add_row(naics_code = code, end_use = "Other Facility Support", .after = row+9) %>% 
    add_row(naics_code = code, end_use = "Onsite Transportation", .after = row+10) %>% 
    add_row(naics_code = code, end_use = "Conventional Electricity Generation", .after = row+11) %>% 
    add_row(naics_code = code, end_use = "Other Nonprocess Use", .after = row+12) %>% 
    add_row(naics_code = code, end_use = "End Use Not Reported", .after = row+13) 
  row = row + 15 #length(end_uses)
}
```

```{r}
scaled_coeffs = full_coeffs
back_count = 0
for (row in 2:nrow(scaled_coeffs)) {
  #print(row)
  back_count = back_count + 1
  #print(back_count)
  if (back_count == 15) {
    back_count = 0
  }
  if (scaled_coeffs$end_use[row] != "TOTAL FUEL CONSUMPTION") {
    for (col in colnames(scaled_coeffs)) {
      if (substr(col, 1, 4) == "elec") {
        scaled_coeffs[row, col] = 
          scaled_coeffs[row - back_count, col] * 
          mecs5_4_wrangled[row, "net_electricity_demand"] /
          mecs5_4_wrangled[row - back_count, "net_electricity_demand"]
        #print(col)
      } else if (substr(col, 1, 2) == "ng") {
        scaled_coeffs[row, col] = 
          scaled_coeffs[row - back_count, col] * 
          mecs5_4_wrangled[row, "nat_gas"] /
          mecs5_4_wrangled[row - back_count, "nat_gas"]
      }
    }
  }
  
}

```

```{r}
# export coeffs to excel file
mecs_other_wrangled = scaled_coeffs
write_xlsx(mecs_other_wrangled, here("outputs", "mecs_other_wrangled.xlsx"))
```

